---
title: "DSeq"
author: "amjad"
date: "03/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(DESeq2)
```
Read in the files

```{r}
counts_table <- read.table("data/obds_countstable.tsv.gz", header = TRUE, row.names = 1)
```


```{r}
sample_table <- read.table("data/obds_sampletable.tsv", header = TRUE, row.names = 1, sep = "\t")
```

Now convert them into a format for generating DSEqData set - need to use the command :
##DESeqDataSetFromMatrix(countData, colData, design, tidy = FALSE,
##ignoreRank = FALSE, ...)
Convert df to matrix
coldata = sample table

First step to using this command:
countData - making a matrix for DESeqDataSetFromMatrix
```{r}
counts_table <- as.matrix(counts_table)
View(counts_table)
```
We need to run the replicate number from the sample table row sample title
```{r}
sample_table <- sample_table %>% separate(sample_title, into = c("egr", "ko", "cell", "replicate"), sep = "_" )%>%
unite("egr_ko_cell", egr, ko, cell, sep = "_")
```

Need to make the egr_ko_cell row into a factor and set the levels

```{r}
sample_table$egr_ko_cell <- factor(sample_table$egr_ko_cell, levels = c("Egr2/3_DKO_CD8", "Egr2/3_DKO_CD4", "Egr2_Kin_CD8", "Egr2_Kin_CD4"))
```

```{r}
dds <- DESeqDataSetFromMatrix(counts_table, sample_table, ~egr_ko_cell)

```
##Access the design formula, counts matrix and sample info from the dds

Design formula
```{r}
dds
show(dds)
str(dds)
head(assays(dds)$count) #Is the counts matrix
assays(dds)
```
Show the design formula. 
```{r}
colData(dds)
```
#Calculate the size factors for each sample - estimatesizefactors
## S4 method for signature 'DESeqDataSet'
estimateSizeFactors(object, type = c("ratio",
  "poscounts", "iterate"), locfunc = stats::median, geoMeans,
  controlGenes, normMatrix, quiet = FALSE)
#Put the size factors into the previous dds object
This is where we are normalising for library size, RNA composition etc. 
```{r}
dds <- estimateSizeFactors(dds)
sizeFactors(dds)
```
Generate a bar plot of the size factors for each sample, coloured by
condition/group.
Make a new df with all this info

frame(..., row.names = NULL, check.rows = FALSE,
           check.names = TRUE, fix.empty.names = TRUE,
           stringsAsFactors = default.stringsAsFactors())
```{r}
row.names(colData(dds))

plot_table <- data.frame(sample = row.names(colData(dds)), sizeFactors(dds), colData(dds)$egr_ko_cell)
condition=colData(dds)$egr_ko_cell
View(plot_table)
```


```{r}
library(ggplot2)
ggplot(plot_table, aes(x = sample, y = sizeFactors.dds., fill=condition))+
    geom_col()+
    theme(axis.text.x = element_text(angle = 30, hjust = 1))
```
Obtain dispersion estimates for each gene – estimateDispersions()
Ø Plot the per-gene dispersion estimates (DESeq2 has a helper function for
this) plotDispEsts(dds)

## S4 method for signature 'DESeqDataSet'
estimateDispersions(object,
  fitType = c("parametric", "local", "mean"), maxit = 100,
  useCR = TRUE, weightThreshold = 0.01, quiet = FALSE,
  modelMatrix = NULL, minmu = 0.5)

```{r}
dds <- estimateDispersions(dds)
head(assays(dds)$mu)
head(dispersions(dds))

plotDispEsts(dds)
```


Perform the Wald test – nbinomWaldTest()
```{r}
dds <- nbinomWaldTest(dds)
```
Use the DESeq() function to perform steps 5-7 (Size factors, dispersion, GLM/WALD) in one go.
We will make a remake the object
```{r}
dds <- DESeqDataSetFromMatrix(counts_table, sample_table, ~egr_ko_cell)
dds <- DESeq(dds)

```
?Access the coefficients of the NB GLM
```{r}
coef_output <- coef(dds)
head(coef_output)
```
10. Access the results table for the comparison between CD8+ and
CD4+ T cells from Egr2/3 DKO mice
Ø Table should contain log2 fold changes and p-values for all genes
```{r}
View(as.data.frame(colData(dds)))
p_values <- results(dds, c("egr_ko_cell","Egr2/3_DKO_CD8","Egr2/3_DKO_CD4"))
p_values_as_df <- as.data.frame(p_values)
```

11. Plot a histogram of the raw and BH-adjusted p-values – do they look
as expected?

First we will do a padjusted graph
```{r}
p_adj_histogram <- ggplot(p_values_as_df, aes(x=padj))+
    geom_histogram()
p_adj_histogram
```

```{r}
p_histogram <- ggplot(p_values_as_df, aes(x=pvalue))+
    geom_histogram()
p_histogram
```
COmbine the above 2 plots
```{r}
library(cowplot)
plot_grid(p_histogram, p_adj_histogram)
```
12. Generate an MA plot of the log2 FC values for all genes
Use th eplot MA function for this.
MA-plot from base means and log fold changes

```{r}
plotMA(p_values)
```
Each point is a gene on this graph (mean expresison of cd4 and log2fold of the cd4. 
This is the cd4 double knockout vs the cd8 double knockouts
The blue values are the more relavent


13. Shrink the log2 FC values using the normal, apeglm and ashr
methods
```{r}
dds_normal <- lfcShrink(dds, contrast = c("egr_ko_cell","Egr2/3_DKO_CD8","Egr2/3_DKO_CD4"), type = "normal")
resultsNames(dds) #

dds_apeglm <- lfcShrink(dds, coef =  c("egr_ko_cell_Egr2.3_DKO_CD4_vs_Egr2.3_DKO_CD8"), type = "apeglm")
dds_ashr <- lfcShrink(dds, contrast = c("egr_ko_cell","Egr2/3_DKO_CD8","Egr2/3_DKO_CD4"), type = "ashr")

plot_grid(dds_apeglm, dds_ashr, dds_ashr)
```

```{r}
plotMA(dds_normal)
p1 <- recordPlot()    # save current plot in R variable

plotMA(dds_apeglm)
p2 <- recordPlot()

plotMA(dds_ashr)
p3 <- recordPlot()

plot_grid(p1, p2, p3, ncol = 3)
```



Generate a results table (one shrinkage method) containing mgi
symbols
Ø Use the org.Mm.eg.db package
Ø Remove all genes with a padj of NA
Ø How many Ensembl IDs are not assigned an mgi symbol and how many mgi
symbols are duplicated?
```{r}
library(org.Mm.eg.db)
#Use one of the tables with the shrunken p-values
View(dds_apeglm)
dds_apeglem_res <- as.data.frame(dds_apeglm)
#ABove gives us the the table. 
```

Now remove NAs from padj value. So first select were padj = NA with is.na

```{r}
dds_apeglem_res_na <- dds_apeglem_res %>% filter(!is.na(padj))
dim(dds_apeglem_res_na)
dim(dds_apeglem_res)
```
Use the mapIds function
#mapIds(x, keys, column, keytype, ..., multiVals)
# x is the database of IDs the function will look into org.Mm.eg.db 
#Keys is the ensemble id you wan to change i.e. rownames
```{r}
keytypes(org.Mm.eg.db) # our keytypes are out ensemble
#we have the ensemble keytypes in the dataframe
# we want to link this to gene name 'SYMBOL'

temp <- mapIds(org.Mm.eg.db, rownames(dds_apeglem_res_na), column = 'SYMBOL', keytype = 'ENSEMBL')
head(temp)

#Next step is to make a data frame
dds_apeglem_res_na_geneid <- as.data.frame(temp)
View(dds_apeglem_res_na_geneid)

#Make a new column in this data.frame with both of our last two table and then join htem together
#put the row 
dds_apeglem_res_na_geneid$ENSEMBL_ID <- rownames(dds_apeglem_res_na_geneid)
dds_apeglem_res_na$ENSEMBL_ID <- rownames(dds_apeglem_res_na_geneid)
View(dds_apeglem_res_na)
View(dds_apeglem_res_na_geneid)
#Use left join to merge the two tables
dds_apeglem_res_final <- left_join(dds_apeglem_res_na_geneid, dds_apeglem_res_na, by = "ENSEMBL_ID")
View(dds_apeglem_res_final)
```

16. Write the results to a csv file
```{r}
write.csv(dds_apeglem_res_final, file = "DE_Results.csv")
```

17. Filter the results table for padj <0.05 and absolute log2 FC >1
```{r}
dds_apeglem_padjlog2 <- dds_apeglem_res_final %>%
  filter(log2FoldChange >1 & padj <0.05)
View(dds_apeglem_padjlog2)
```

18. 



