---
title: "R_droplet"
author: "amjad"
date: "10/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



renv::init()
BiocManager::install("BiocNeighbours")

After installation of DropletUtils
#BiocManager::install("DropletUtils")


```{r}
library(DropletUtils)
```

```{r}
sce <- read10xCounts("data/pbmc_1k_v3_raw", sample.names = "PBMC")
sce

View(as.data.frame(colData(sce)))
```
Remove empty droplets 

```{r}
head(assay(sce, "counts"))

#Remove droplets without any count at all
sums <- colSums(assay(sce, "counts"))
sce <- sce[, sums >0]
sce
```


```{r}
emptyDroplets <- DropletUtils::emptyDrops(assay(sce, "counts"))
```
#FDR - false discovery rate. Less than 0.01pval considered to be a cell
```{r}
is.cell <- emptyDroplets$FDR <= 0.01
sum(is.cell, na.rm=TRUE)

table(is.cell)
```


```{r}
df <- data.frame(librarysize = colSums(assay(sce, "counts")), emptydroplet = is.cell)

#install.packages("ggplot2")
library(ggplot2)    
ggplot(df, aes(x = emptydroplet, y = librarysize))+
geom_point()
```


```{r}
head(which(is.cell))

# remove FALSE and NA droplets from raw data
sce <- sce[,which(is.cell)]
```

```{r}
sce_filtered <- read10xCounts("data/pbmc_1k_v3_filtered", sample.names = "PBMC")
sce_filtered

# Filtered has 1222 cells, we have 1206
```
```{r}
# compare cells in raw and filtered
barcodes_raw <- colData(sce)$Barcode
barcodes_filtered <- colData(sce_filtered)$Barcode

length(intersect(barcodes_raw, barcodes_filtered))
```


Remove doublets

```{r}
#install.packages("BiocManager")
#BiocManager::install("scDbFinder")
#BiocManager::install("scDblFinder")
install.packages(scDblFinder)
library(scDblFinder)

sce <- scDblFinder(sce)

View(as.data.frame(colData(sce)))
```

Draw a plot comparing doublets with other doublets
```{r}
df2 <- data.frame(librarysize= colSums(assay(sce, "counts")), colData(sce)$scDblFinder.class)
View(df2)

plot2 <- ggplot(df2, aes(x = doublets, y = librarysize))+
  geom_violin()+
  geom_point()
plot2
                  
```

```{r}
head(colData(sce)$scDblFinder.class == "singlet")

sce <- sce[,colData(sce)$scDblFinder.class == "singlet"]
sce
```


```{r}
BiocManager:Install("celda")
library(celda)

sce <- decontX(sce)
assays(sce)$decontXcounts
```

Visualise the UMAP layout created by decontC; use plotDimReduceCluster()
```{r}
#View(as.data.frame(colData(sce)))

#reducedDimNames(sce)

#plotDimReduceCluster(assays(sce)$decontX_clusters, dim1 = reducedDim(sce)$DecontX_UMAP_1, dim2 = reducedDim(sce)$DecontX_UMAP_2)

reducedDim(sce)

class(sce)

plotDimReduceCluster(colData(sce)$decontX_clusters, dim1 = reducedDim(sce)[,"DecontX_UMAP_1"], dim2= reducedDim(sce)[,"DecontX_UMAP_2"])
```

Visualize the percentage of contamination estimated by decontX in each cell. Use the
plotDecontXContamination() function

```{r}
ggplot_object <- plotDecontXContamination(sce)
plotDecontXContamination(sce)

```

Visualize the log-normalised expression of marker genes on a UMAP layout faceted by gene; use
plotDimReduceFeature()

```{r}
reducedDimNames(sce)

head(rownames(sce))

plotDimReduceFeature(as.matrix(assays(sce)$counts), features = "ENSG00000243485", dim1 = reducedDim(sce)[,"DecontX_UMAP_1"], dim2= reducedDim(sce)[,"DecontX_UMAP_2"])
```



