---
title: "R_MLexercise"
author: "amjad"
date: "02/03/2021"
output: html_document
---
```{r}
library(caret)
```

Get familiar with the data set. What are rows? What are columns?

```{r}
data(iris)
View(iris)
```

Use the caret::featurePlot() function to visualise species and measurements. Try the different options
for the plot argument.

featurePlot(
  x,
  y,
  plot = if (is.factor(y)) "strip" else "scatter",
  labels = c("Feature", ""),

```{r}
caret::featurePlot(x=iris[,1:4],y=iris$Species, plot='density', auto.key = list(columns = 3))

```
Partition the data set into one training and one test set. Make sure the two sets are balanced with
respect to Species .

```{r}
set.seed(998)
inTraining <- createDataPartition(iris$Species, p = .75, list = FALSE)
training <- iris[ inTraining,]
testing <- iris[-inTraining,]
```

```{r}
library(e1071)
fitControl <- trainControl(method = "boot",number = 5)
knnFit <- train(Species ~ ., data = training, method = "knn", trControl = fitControl, tuneGrid = data.frame(k = c(1,2,5,10,20)))
ggplot(knnFit)
```
```{r}
library(kernlab)
fitControl <- trainControl(method = "boot",number = 5)
GPRFit <- train(Species ~ ., data = training, method = 'gaussprLinear', trControl = fitControl)
ggplot(GPRFit)
```


```{r}
knnPred <- predict(knnFit, newdata = testing)
```

```{r}
confusionMatrix(data = knnPred, testing$Species)$table
```

```{r}
confusionMatrix(data = knnPred, testing$Species)$overall["Accuracy"]

```




