---
title: "R_MLearning"
author: "amjad"
date: "02/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Fit a linear machine learning model to chickweight to measure the effect of weight and diet. 

Which diet leads to the fastest increase in body weight?
How much does weight increase per unit of Time for the top diet?
Does the top diet drive an increase in body weight that is signicantly faster than the next best diet?


```{r}
library(tidyverse)
data("ChickWeight")
```

There are 4 diets. 
```{r}
View(ChickWeight)
lm(weight ~ Time + Diet, ChickWeight)

```


```{r}
ggplot(ChickWeight, aes(Time, weight, color=Diet)) +
    geom_point()+
    stat_smooth(method='lm')
```
Below we are predicting wieght over time without any effect of diet. 
So at time 0 they are 27.467 and for every unit of time they gain 8.803g. Unit of time = 1 day. 

```{r}
lm(weight ~ Time, ChickWeight)
```
Plot above on a graph (Weight vs time). We can see 0 - 27.467

```{r}
ggplot(ChickWeight, aes(Time, weight)) +
    geom_point()+
    stat_smooth(method='lm')
```
Now calculate if they will gain 8.8g i.e. intercept. So we use geom_abline and use our values. This should fit a line on the ggplot line

```{r}
ggplot(ChickWeight, aes(Time, weight)) +
    geom_point()+
    stat_smooth(method='lm')
    geom_abline(slope=8.803, intercept=27.467)
```

Now check weight with diet
```{r}
lm(weight ~ Diet, ChickWeight)
```
Interpretation of above
A chickweight on diet 1 is on average 102.65g irrespective of time, at any time between day 1 to 25. 
A chickweight on diet 2 is on average 20g higher than diet 1. 
A chickweight on diet 3 is on average 40g higher than diet 1.

Weight vs diet graph
It uses diet one as the intercept

```{r}
ggplot(ChickWeight, aes(Diet, weight)) +
    geom_boxplot()+
    geom_point()
```
Starting to build model

```{r}
lm(formula = weight ~ Time * Diet, data = ChickWeight)
```
30.9310 is interceot for diet 1, diet 2 on average has 2.2g less at time 0. 
-12.68 is intercept for diet 3 below diet 1
Time:Diet3 is the change in weight relative to diet 1
Time:Diet is effectively the slope of gain. 

```{r}
ggplot(ChickWeight, aes(Time, weight, color=Diet)) +
    geom_point()+
    stat_smooth(method='lm')+
```
Diet 3 leads to the fastest incraese in body weight (Time:Diet3)

How much doe sthe weight incraese per unit of TIme for the top diet?
- Time diet 1(6.8418) + time diet 3 (4.5811) = 11.4229
Time diet 1 is the base line incraese for all diets. Their values are in addition to this. 

Intercept for diet 3 = intercept diet 1 - diet 3 (30 - 12)

```{r}
ggplot(ChickWeight, aes(Time, weight, color=Diet)) +
    geom_point()+
    stat_smooth(method='lm')+
    geom_abline(slope=11.4229, intercept = 30.9310 - 12.6807)
```
```{r}
chickweight_lm_fit <-  lm(formula = weight ~ Time * Diet, data = ChickWeight)
summary(chickweight_lm_fit)
```

Does the top diet drive an incraes in body weight that is significantly faster than the next best diet?


Now compare to diets to diet 4. NOte the reference level is only compared to 0. 
```{r}
ChickWeight$Diet <- relevel(ChickWeight$Diet, "4")
chickweight_lm_fit <-  lm(formula = weight ~ Time * Diet, data = ChickWeight)
summary(chickweight_lm_fit)
```

(Intercept)  30.7921     5.9209   5.201 2.77e-07 ***
Time          9.7144     0.4670  20.802  < 2e-16 ***

ABOVE TWO ARE COMPARING DIET 4 TO 0
THE REMAINING LINES WITHIN THE DATA COMPARE DIETS 1-3 WITH 4. 

```{r}
predict(chickweight_lm_fit, tibble(Time = 5, Diet = '3'))
```

To calculate the weight of chick 3 on day 5
predict(chickweight_lm_fit, tibble(Time = 5, Diet = '3'))

