---
title: "Integration"
author: "amjad"
date: "11/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import the two data sets
# How to load RDS file
```{r}
pbmcv2 <- readRDS("Data/pbmcv2_filtered.rds")
pbmcv3 <- readRDS("Data/pbmcv3_filtered.rds")

```
#What are the two file types
```{r}
pbmcv2
```

```{r}
pbmcv3
```
Note they have different features (genes i.e. 13070 vs 15246)

#Create a seurat object for each of the above 2 files. We do not need to do this as both are seurat objects. So we can merge 


```{r}
pbmc.combined <- merge(x = pbmcv2, y = pbmcv3, add.cell.ids = c("v2", "v3"), project = "PBMC")
pbmc.combined
```

```{r}
union_genes <- unique(c(rownames(pbmcv2), rownames(pbmcv3)))
head(union_genes)
identical(union_genes, rownames(pbmc.combined))
```
#3. On the merged Seurat object, perform the general steps of a Seurat analysis workflow:
- Normalisation
- Finding variable genes
- Scaling
- PCA
- Clustering
- UMAP
4. Visualise the identified clusters on the UMAP
5. Colour the cells in the UMAP by the original dataset from which they came i.e. v2 or v3
- Do you see any batch effects?

Use the seural vignetes
```{r}
pbmc.combined[["percent.mt"]] <- PercentageFeatureSet(pbmc.combined, pattern = "^MT-")
pbmc.combined
```


```{r}
VlnPlot(pbmc.combined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```

```{r}
plot1 <- FeatureScatter(pbmc.combined, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(pbmc.combined, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 + plot2
```
Normalise the data
```{r}
pbmc.combined <- NormalizeData(pbmc.combined, normalization.method = "LogNormalize", scale.factor = 10000)
```

Identification of highly variable features
```{r}
pbmc <- FindVariableFeatures(pbmc.combined, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(pbmc), 10)

# plot variable features with and without labels
plot3 <- VariableFeaturePlot(pbmc)
plot4 <- LabelPoints(plot = plot3, points = top10, repel = TRUE)
plot3 + plot4
```

Scale the data
```{r}
all.genes <- rownames(pbmc)
pbmc <- ScaleData(pbmc, features = all.genes)
```



```{r}
pbmc <- RunPCA(pbmc, features = VariableFeatures(object = pbmc))
```

```{r}
print(pbmc[["pca"]], dims = 1:5, nfeatures = 5)
```

```{r}
VizDimLoadings(pbmc, dims = 1:2, reduction = "pca")
```

```{r}
DimPlot(pbmc, reduction = "pca")
```
Kevins code for bonus exercise

```{r}
library(ggplot2)
df <- data.frame(
pbmc.combined@reductions$pca@cell.embeddings[, 1:2],
ident = pbmc.combined@meta.data$orig.ident
)
df <- df[sample(nrow(df)), ]
dimplot_ggplot2 <- ggplot(df, aes(PC_1, PC_2, color = ident)) +
geom_point() +
theme_classic()
dimplot_ggplot2
dimplot_seurat + dimplot_ggplot2
```


```{r}
DimHeatmap(pbmc, dims = 1, cells = 500, balanced = TRUE)
```

```{r}
DimHeatmap(pbmc, dims = 1:9, cells = 500, balanced = TRUE)
```
Determine dimensionality
```{r}
pbmc <- JackStraw(pbmc, num.replicate = 100)
pbmc <- ScoreJackStraw(pbmc, dims = 1:20)
```
```{r}
JackStrawPlot(pbmc, dims = 1:15)
```

```{r}
ElbowPlot(pbmc)
```

Cluster the cells

```{r}
pbmc <- FindNeighbors(pbmc, dims = 1:10)
pbmc <- FindClusters(pbmc, resolution = 0.5)
```

```{r}
# Look at cluster IDs of the first 5 cells
head(Idents(pbmc), 5)
```

##Run non-linear dimensional reduction (UMAP/tSNE)

```{r}
pbmc <- RunUMAP(pbmc, dims = 1:10)
```
4. Visualise the identified clusters on the UMAP
5. Colour the cells in the UMAP by the original dataset from which they came i.e. v2 or v3
- Do you see any batch effects?
```{r}
DimPlot(pbmc, reduction = "umap", label = TRUE)
```


#color by batch v2 and v3

```{r}
#group by allows us to color by batch/group
colnames(pbmc@meta.data)
DimPlot(pbmc, reduction = "umap", group.by = 'orig.ident', label = TRUE)

```
Above graph shows there is batch effect. So naive analysis can't continue any further. 



#EXERCISE 2

We need to split into list first
https://satijalab.org/seurat/archive/v3.0/integration.html


```{r}
pbmc.list <-  SplitObject(pbmc, split.by = "orig.ident")
names(pbmc.list)

features <- SelectIntegrationFeatures((object.list = pbmc.list))
anchors <- FindIntegrationAnchors(object.list = pbmc.list, anchor.features = features)
anchors

pbmc.integrated <- IntegrateData(anchorset = anchors)

```
```{r}
library(Seurat)
library(SeuratData)
library(patchwork)

DefaultAssay(pbmc.integrated) <- "integrated"

pbmc.integrated <- ScaleData(pbmc.integrated, npcs=30, verbose = FALSE)
pbmc.integrated <- RunPCA(pbmc.integrated, features = VariableFeatures(object = pbmc.integrated))

```

```{r}
library(ggplot2)
library(cowplot)
```


```{r}
# Run the standard workflow for visualization and clustering
pbmc.integrated <- ScaleData(pbmc.integrated, verbose = FALSE)
pbmc.integrated <- RunPCA(pbmc.integrated, npcs = 30, verbose = FALSE)
pbmc.integrated <- RunUMAP(pbmc.integrated, reduction = "pca", dims = 1:30)
p1 <- DimPlot(pbmc.integrated, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(pbmc.integrated, reduction = "umap", group.by = "seurat_clusters", label = TRUE, 
    repel = TRUE) + NoLegend()
plot_grid(p1, p2)
```
https://satijalab.org/seurat/articles/integration_introduction.html

```{r}
pbmc.integrated <- RunUMAP(pbmc.integrated, reduction = "pca", dims = 1:20)
pbmc.integrated <- FindNeighbors(pbmc.integrated, reduction = "pca", dims = 1:20)
pbmc.integrated <- FindClusters(pbmc.integrated, resolution = 0.5)
```

```{r}
p1 <- DimPlot(pbmc.integrated, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(pbmc.integrated, reduction = "umap", label = TRUE, repel = TRUE)
p1 + p2
```

```{r}
DefaultAssay(pbmc.integrated) <- "RNA"
markers <- FindConservedMarkers(pbmc.integrated, ident.1 = 6, grouping.var = "orig.ident", verbose = FALSE)
head(markers)
```









