--title: "R_dimensionality"
author: "amjad"
date: "01/03/2021"
output: html_document
---


```{r}
logcounts <- read.csv('data/logcounts.csv', row.names = 1)
View(logcounts)
```

```{r}
str(logcounts)
```
The class command will tell us the dataframe or matric, but not the contents

```{r}
logcounts_matrix <- data.matrix(logcounts, rownames.force = NA)
class(logcounts_matrix)
```

Question 2. Perform PCA

```{r}
logcounts_pca <- prcomp(logcounts_matrix, centre = TRUE, scale. = FALSE)
summary(logcounts_pca)
```
```{r}
str(logcounts_pca)
```
Looking at above the rows and columns are the wrong ways. 
We need clustering on samples not on genes
i.e. rotation needs to GENEID:PC1 and not cell id 
So we have to transpose the data (original marix). The rerun the PCA

```{r}
transposed_logcounts <- t(logcounts_matrix)
str(transposed_logcounts)
```
```{r}
transposed_logcounts_pca <- prcomp(transposed_logcounts, centre = TRUE, scale. = FALSE)
str(transposed_logcounts_pca)
```
Now the rotation has GENEID:PC1, etc
We ca view the new matrix coordinates using the the View(variable_name$x)

```{r}
View(transposed_logcounts_pca$x)
```

```{r}
renv::restore()
```


3. Draw a scatterplot of the top two principle components and color by metadata in cell_metadata.csv. Which experimental factors show the largest variance. 
Sample = Sample - needs to be an exact match (Sample 1 is row from tabel 1 and sample 2 is row from table 2)

```{r}
cell_metadata <- read.csv("data/cell_metadata.csv")

logcounts_pca_df <- as.data.frame(transposed_logcounts_pca$x[,1:2])

#create a smaple with row names
logcounts_pca_df$Sample <- rownames(logcounts_pca_df)
View(logcounts_pca_df)

library(tidyverse)
cell_counts <- logcounts_pca_df %>%
    left_join(cell_metadata, by = c("Sample" = "Sample"))
View(cell_counts)
```



```{r}
library(ggplot2)
plot1 <- ggplot(data = cell_counts, aes(x=PC1, y=PC2, col = Infection))+
    geom_point()
plot1
```
Draw a density plot, faceted by time and infection, colored by status, to inspect the variance of each
experimental factor on the top principal component.
1/ density plot
2/ facet time
3/ facet infection

1/...Below

```{r}
plot2 <-ggplot(data = cell_counts, aes(x=PC1, col = Infection))+
    geom_density()
plot2

```
Next step with facet time and infection

```{r}
plot2b <-ggplot(data = cell_counts, aes(x=PC1, col = Status))+
    geom_density() +
    facet_grid(rows =vars(Infection), cols = vars(Time)) 
plot2b
```
Draw a bar plot of variance explained by each principal component. How many principal components
would you keep for downstream analyses?

```{r}
variance_table <- data.frame(variance =(transposed_logcounts_pca$sdev)^2)
variance_table$fraction <- variance_table$variance/sum(variance_table$variance)
PC_numbers <- seq(1, nrow(variance_table))
variance_table$PC <- PC_numbers
temp <- variance_table %>%
    select(PC, variance, fraction)
View(variance_table)

```



```{r}
plot3 <- ggplot(variance_table, aes(x = PC, y=fraction))+
    geom_col() +
    xlim(0, 100)
plot3

```

Find the names of the top genes associated with the top principal components.

```{r}
variance_table$cum_sum <- cumsum(variance_table$fraction)
View(variance_table)

Plot4 <- ggplot(variance_table, aes(x = PC, y = cum_sum))+
    geom_col() +
    xlim(0, 100)
Plot4
```

Line plot of above


```{r}
Plot4b <- ggplot(variance_table, aes(x = PC, y = cum_sum))+
    geom_line() +
    xlim(0, 100)
Plot4b
```
Visualise the gene expression value of the gene with the highest loading for PC1 as color on a scatter
plot of PC1 and PC2. What do you observe?

```{r}
gene_influence <- as.data.frame(logcounts_pca$rotation)
gene_influence$gene_ID <- rownames(gene_influence)
gene_influence <- gene_influence %>%
    arrange(desc(PC1)) %>%
    select(gene_ID, PC1)
View(gene_influence)
```


```{r}

```

