---
title: "Thursday"
author: "amjad"
date: "25/02/2021"
output: html_document
---

## R Markdown

```{r}
getwd()
counts_table <- read_tsv("data/obds_countstable.tsv.gz")
View(counts_table)
```

```{r}
sample_table <- read_tsv("data/obds_sampletable.tsv")
View(sample_table)
```
```{r}
counts_table_tidy <- counts_table%>%
    pivot_longer(-Geneid, names_to = "samples", values_to = "count")
View(counts_table_tidy)
```

With the above we have converted the coutns table in to three columns. All Geneids, then column 2=sample, column 3=count, for all 12 ids

Now join geneinfo to get mgi_symbol
Use biomart

```{r}
library(biomaRt)
#Download the database. We are loading the data base into a variable called ensembl
ensembl <- useMart("ensembl")
#Get a list of datasets to work out which one to use. I.e. we are lisiting all the samples. 
listDatasets(ensembl)
#load the mosue database. So we are making the databse again but only with the mouse info we have selected. 
ensembl <- useMart("ensembl",dataset="mmusculus_gene_ensembl")
#See which attributes/ensemble IDs are avaialble
listAttributes(ensembl)
#So we want to macth the ensemble_gene_id (form our table) to mgi_symbol
# now we need to write a query function
# retrive both to know the pairs
gene_names <- getBM(c("ensembl_gene_id","mgi_symbol"), filters = "ensembl_gene_id", values = counts_table$Geneid, mart = ensembl)
View(gene_names)

```
Now join with the other table i.e counts_table_tidy with gene_names
First check for duplicates. 

```{r}
length(unique(gene_names$ensembl_gene_id))
nrow(gene_names)
```

Join the tables now, make sure the columns are named the same with the 'by' function

```{r}
counts_table_tidy <- counts_table_tidy%>%
    left_join(gene_names, by = c("Geneid" = "ensembl_gene_id"))
#Are there any unanmed genes
sum(is.na(counts_table_tidy$mgi_symbol))/12

```
Tidy the metadata file


```{r}
sample_table <- sample_table %>%
  separate(sample_title, c("gene_name","gene_condition", "cell_type","replicate_number"), sep = "_")

sample_table <- sample_table %>% 
  unite(col = genotype, gene_name, gene_condition, sep = "_")

sample_table <- sample_table %>%
  dplyr::select(-species, -library_layout)
View(sample_table)
```

Add metadata to table with counts and gene info
```{r}
View(counts_table_tidy)
View(sample_table)
counts_table_tidy <- counts_table_tidy %>% 
    left_join(sample_table, by = c("samples" = "Sample_accession"))
```

We express counts per sample as the number of counts per sample. How many cpm is there for each of these genes. 
Use mutate function. Mutate(New coloum called cpm containing....)

```{r}
counts_table_tidy <- counts_table_tidy%>%
  mutate(CPM = count*1000000/read_count)
```

Also calculate log2(CPM +1)
```{r}
counts_table_tidy <- counts_table_tidy%>%
  mutate(log_transformed = log2(CPM + 1))
```

Remove some rows (CPM, log transformed and read count). We are removing them as we have accidently been given the answer
To select the right columns we have to use dplyr
```{r}
counts_table_tidy <- counts_table_tidy %>%
  dplyr::select(Geneid:replicate_number)
View(temp)
```

```{r}
counts_table_tidy <- counts_table_tidy %>%
  group_by(samples)%>%
  mutate(total_counts = sum(count))
View(temp)

#Check if ther are 12 unique counts
unique(temp$total_counts)
```
```{r}
counts_table_tidy <- counts_table_tidy %>%
  mutate(CPM = count*1000000/total_counts)
```

```{r}
counts_table_tidy <- counts_table_tidy %>%
  mutate(log_transformed = log2(CPM + 1))

```


```{r}
plotting_table <- counts_table_tidy %>%
  group_by(samples) %>%
  summarise(total_counts = sum(count))
plotting_table

plot1 <- ggplot(plotting_table, aes(x = samples, y = total_counts)) +
                geom_col()+
theme(axis.title = element_text(size = 8, face = 'bold'), axis.text.x = element_text(size = 8, angle = 30, hjust = 1, color = 'red'))
plot1
```
Summarise by gene
a) How many genes have no counts for any sample?
```{r}
temp <- counts_table_tidy %>%
  group_by(mgi_symbol) %>%
  summarise(total_counts_per_gene = sum(count)) %>%
  filter(total_counts_per_gene == 0) %>%
  tally()

```
b) Draw a density plot of log2(CPM + 1) for all genes
```{r}
plot2 <- plot1 <- ggplot(counts_table_tidy, aes(x = log_transformed, col = samples))+
  geom_density()+
  theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 1, color = 'red'))
plot2
  
```
Filter out genes that have low expression in 3 or fewer samples

```{r}
filtered_genes <- counts_table_tidy %>%
  group_by(Geneid) %>%
  filter(CPM >= 0.5) %>% #filter out genes with <0.5 count
  tally() %>%
  filter(n > 3)%>%
  pull(Geneid)
filtered_genes
```

Make a density plot of log2(CPM + 1) with the ltered data

```{r}
filtered_counts_table <- counts_table_tidy %>%
  filter(Geneid %in% filtered_genes)
nrow(filtered_counts_table)
nrow(counts_table_tidy)
```

```{r}
temp <- counts_table_tidy %>%
  group_by(Geneid) %>%
  filter(sum(CPM>=0.5)>3)

```


```{r}
plot3 <- ggplot(filtered_counts_table, aes(x = log_transformed, col = samples))+
  geom_density()+
  theme(axis.text.x = element_text(size = 10, angle = 90, hjust = 1, color = 'red'))
plot3
  
```

