---
title: "ggplot2_exercise"
author: "amjad"
date: "24/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```
```{r}
ggplot(diamonds)
```
```{r}
ggplot(diamonds, aes(x = carat, y = price))
```
```{r}
ggplot(diamonds, aes(x = carat, y = price))+
    geom_point()
```
```{r}
ggplot(diamonds, aes(x = carat, y = price, color=cut))+
    geom_point()
```
Follow the mean of the data point
```{r}
ggplot(diamonds, aes(x = carat, y = price, colour = cut)) +
 geom_point() +
 geom_smooth()

```
Make into a single line. Uses a method called gam to fit the line. 
```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
 geom_point(aes(colour = cut)) +
 geom_smooth()

```
Edit the theme i.e. axis tick/text/background. So we have to make a new function for this. 
```{r}
base_plot +
 labs(title = "Diamond price and carat values",
 x = "Carat", y = "Price") +
 theme(axis.title = element_text(size = 16, face = "bold"),
 axis.text = element_text(size = 14),
 plot.title = element_text(hjust = 0.5, size = 20))


```
```{r}
ChickWeight
```
Working with new dataset 'ChickWeight'
Plot a histogram
#geom_histogram() - histogram, requires x (continuous)

ggplot(diamonds, aes(x = carat, y = price, colour = cut)) +
 geom_point() +
 geom_smooth()
```{r}
ggplot(ChickWeight,aes(x = weight) ) + 
geom_histogram()
```
Change fill and border
```{r}
ggplot(ChickWeight,aes(x = weight) ) + 
geom_histogram(fill = "grey", color = "black")
```
```{r}
ggplot(ChickWeight,aes(x = weight) ) + 
geom_histogram(fill = "grey", color = "black", binwidth = 100)
```
Above we can change the binwidth or the no. of bins. Cant have both.
Change the label and font size of the axes.

```{r}
ggplot(ChickWeight,aes(x = weight) ) + 
geom_histogram(fill = "grey", color = "black", binwidth = 10) +
labs(x = 'chick weight', y = 'Number of Chicks', title = 'Histogram of chick weight', subtitle = 'chickweight data set') +
theme(axis.title = element_text(size = 16, face = 'bold'), axis.text.x = element_text(size = 14, angle = 30, hjust = 0, color = 'red'))
```
Change the y-axis upper limit to an arbitrary number, e.g. 100

```{r}
ggplot(ChickWeight,aes(x = weight) ) + 
geom_histogram(fill = "grey", color = "black", binwidth = 10) +
labs(x = 'chick weight', y = 'Number of Chicks', title = 'Histogram of chick weight', subtitle = 'chickweight data set') +
theme(axis.title = element_text(size = 16, face = 'bold'), axis.text.x = element_text(size = 14, angle = 30, hjust = 0, color = 'red')) + 
    ylim(0,100)
```
SPlit data in to the four diets

```{r}
ggplot(ChickWeight,aes(x = weight) ) + 
geom_histogram(fill = "grey", color = "black", binwidth = 10) +
labs(x = 'chick weight', y = 'Number of Chicks', title = 'Histogram of chick weight', subtitle = 'chickweight data set') +
theme(axis.title = element_text(size = 16, face = 'bold'), axis.text.x = element_text(size = 14, angle = 30, hjust = 0, color = 'red')) + 
    ylim(0,25) +
    facet_wrap(~Diet, nrow = 4)

```
Color each diet in a different color. So we are mapping data to a plot feature
```{r}
ggplot(ChickWeight,aes(x = weight, fill = Diet) ) + 
geom_histogram(color = 'black', binwidth = 5) +
labs(x = 'chick weight', y = 'Number of Chicks', title = 'Histogram of chick weight', subtitle = 'chickweight data set') +
theme(axis.title = element_text(size = 16, face = 'bold'), axis.text.x = element_text(size = 14, angle = 30, hjust = 0, color = 'red')) + 
    ylim(0,25) +
    facet_wrap(~Diet, nrow = 4)
```
Allow y axis to adapt to the range in each diet. Facet_wrap. 

```{r}
ggplot(ChickWeight,aes(x = weight, fill = Diet) ) + 
geom_histogram(color = 'black', binwidth = 5) +
labs(x = 'chick weight', y = 'Number of Chicks', title = 'Histogram of chick weight', subtitle = 'chickweight data set') +
theme(axis.title = element_text(size = 16, face = 'bold'), axis.text.x = element_text(size = 14, angle = 30, hjust = 0, color = 'red')) + 
    facet_wrap(~Diet, ncol = 1)
```
Create a scatter plot of weight (y-axis) over time (x-axis)
Add colour
Add mean trend 
Which diet leads to the highest mean increase in weight?
Collate that last plot with the last histogram in the previous exercise.
Label the panels "A" and "B".

```{r}
ggplot(ChickWeight,aes(x = Time, y = weight, color = Diet) )+
    geom_point() +
    geom_smooth(method=lm)
    

```

```{r}
collate 
```

